\documentclass[]{bytedance_seed}

% single-column: \documentclass[]{bytedance_seed}, 
%Please prioritize using single-column。

% twocolumn: \documentclass[twocolumn]{bytedance_seed}

\usepackage[toc,page,header]{appendix}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{minitoc}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{amsfonts}
\usepackage{amssymb}







%%%%%%%%%%%%%%%%%%%%

\title{VAPO: Efficient and Reliable Reinforcement Learning for Advanced Reasoning Tasks }

% \author[1]{ByteDance Seed}
% \author[1]{Author2}
% \author[1,2,*]{Author3}
% \author[1]{Author4}
% \author[1, \dagger]{Author5}

%论文单位请使用ByteDance Seed
\affiliation[]{ByteDance Seed}
\contribution{Full author list in Contributions}


\abstract{
We present VAPO, \textbf{V}alue-model-based \textbf{A}ugmented Proximal \textbf{P}olicy \textbf{O}ptimization framework for reasoning models., a novel framework tailored for reasoning models within the value-model-based paradigm. Benchmarked  the AIME 2024 dataset, VAPO, built on the Qwen 32B pre-trained model, attains a state-of-the-art score of $\mathbf{60.4}$. In direct comparison under identical experimental settings, VAPO outperforms the previously reported results of DeepSeek-R1-Zero-Qwen-32B and DAPO by more than 10 points. The training process of VAPO stands out for its stability and efficiency. It reaches state-of-the-art performance within a mere 5,000 steps. Moreover, across multiple independent runs, no training crashes occur, underscoring its reliability. This research delves into long chain-of-thought (long-CoT) reasoning using a value-model-based reinforcement learning framework. We pinpoint three key challenges that plague value-model-based methods: value model bias, the presence of heterogeneous sequence lengths, and the sparsity of reward signals. Through systematic design, VAPO offers an integrated solution that effectively alleviates these challenges, enabling enhanced performance in long-CoT reasoning tasks.
}

\date{\today}
\correspondence{Yu Yue at \email{yueyu@bytedance.com}}

% You can add additional info fields as follows 

\begin{document}
\maketitle

%不需要目录就注释掉 注意目录不要和第一页放在一块 要有\newpage
%\newpage
%\tableofcontents
%\newpage

\input{sections/introduction}
\input{sections/relatedwork}
\input{sections/approach}
\input{sections/experiments}
\input{sections/conclusion}

\clearpage

\bibliographystyle{plainnat}
\bibliography{main}

%\clearpage

\end{document}