3D Gaussian Splatting (3DGS) has demonstrated impressive novel view synthesis performance. 
While conventional methods require per-scene optimization, more recently several feed-forward methods have been proposed to generate pixel-aligned Gaussian representations with a learnable network, which are generalizable to different scenes. 
However, these methods simply combine pixel-aligned Gaussians from multiple views as scene representations, thereby leading to artifacts and extra memory cost without fully capturing the relations of Gaussians from different images. 
In this paper, we propose Gaussian Graph Network (GGN) to generate efficient and generalizable Gaussian representations.
Specifically, we construct Gaussian Graphs to model the relations of Gaussian groups from different views.
To support message passing at Gaussian level, we reformulate the basic graph operations over Gaussian representations, enabling each Gaussian to benefit from its connected Gaussian groups with Gaussian feature fusion.
Furthermore, we design a Gaussian pooling layer to aggregate various Gaussian groups for efficient representations.
We conduct experiments on the large-scale RealEstate10K and ACID datasets to demonstrate the efficiency and generalization of our method. Compared to the state-of-the-art methods, our model uses fewer Gaussians and achieves better image quality with higher rendering speed.
 
% 3D Gaussian Splatting (3DGS) has demonstrated impressive novel view synthesis performance, but requires per-scene optimization. Thus, several feed-forward Gaussian Splatting methods have been proposed to generate pixel-aligned Gaussian representations via a learnable network, which can be generalized to different scenes. 
% However, the simple combination of Gaussians from input views achieves superfluous Gaussian representations, for the overlapped regions are covered by similar Gaussians predicted separately from multiple images. This redundancy of Gaussians contributes to ghost in rendering novel view and extra memory cost.
% In view of this challenge, we propose Gaussian Graph Network (GGN) to generate efficient and generalizable Gaussian representations. We build a graph to describe the relations between Gaussians from different views, and extend the definition of linear and pooling layers in graph neural networks to Gaussian domain.  
% Specifically, linear layers update Gaussian features based on the graph structure, while pooling layers merge similar Gaussians to obtain efficient representations.
% Extensive qualitative and quantitative experiments illustrate that our method achieves better rendering quality with fewer Gaussian quantity.